### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form

@@

We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
\begin{equation}
\label{eq:wi_optimization_problem_intro}
\begin{aligned}
\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the functions $w_i$ are the flattest functions, in a least-squares sense, that interpolate the points $x_i$. Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid. We call this interpolation scheme \emph{Poisson interpolation}. The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions. This is illustrated in Figure FIG.

@@

\subsection{Background: discretization}
\label{sec:discretization}

@@

\section{Sampling impulse responses}

We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points. The challenge is to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$.

The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$. Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$. We therefore choose sample points $x_i$ from a region $\Omega_I \subset \Omega$ which is not too close to the boundary. We determine $\Omega_I$ by applying $\mathcal{A}^T$ to a boundary source, then post processing the result.

@@

@@

\paragraph{Discretization} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively. The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
	\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
	\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
	\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
\end{align*}
respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.

\subsection{Choosing sample points}
\label{sec:choosing_sample_points}

@@

\paragraph{Discretization}
In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$. The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$. The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
\begin{equation*}
	\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
\end{equation*}
The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
\begin{equation*}
	\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
\end{equation*}

@@

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness [cite Jacobsen], the weighting function $w_i$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and takes the value $1$ at point $x_i$, and zero at the points $x_j$, $j \neq i$. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. Then in Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added. However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.

It is important to note that optimization problems \eqref{eq:wi_optimization_problem} only has a unique continuous solution when the spatial dimension is $1$, $2$, or $3$. When inverting for parameter fields defined on $4$ or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions. In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.

\begin{thm}
	\label{thm:smoothest_interpolant}
Let $u$ be the function
\begin{equation*}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation*}
for an arbitrary but fixed value of $y$. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
\end{proof}

@@

\begin{thm}
	The weighting functions form a partition of unity.
\end{thm}

@@

\section{Discretized function space operations}
\label{app:discretized_operations}

@@

>>>>

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form

@@

We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
\begin{equation}
\label{eq:wi_optimization_problem_intro}
\begin{aligned}
\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$. Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid. We call this interpolation scheme \emph{Poisson interpolation}. The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions. This is illustrated in Figure FIG.

@@

@@

\section{Choosing sample points}
\label{sec:choosing_sample_points}

We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points. In this section we describe how we choose the points $x_i$. The challenge is to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$.

@@

The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$. Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$. We therefore choose exclude sample points $x_i$ if the ellipsoid associated with that point overlaps the boundary.

@@

\subsection{Greedy point selection algorithm}
\label{sec:greedy_point_selection}

@@

@@

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. Theorem \ref{thm:smoothest_interpolant} relies on some straightforward prerequisites that we present in Definition \ref{def:stiffness_mass}, Proposition \ref{prop:equivalent_writing_of_weighting_optimization}, and Corollary \ref{cor:linearity_of_solution_operator}. In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.

In Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added. However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.

It is important to note that functions in $H^2$ only have pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions. In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.

\begin{defn}
	\label{def:stiffness_mass}
	We define $\mathcal{K}:H^2(\Omega)\rightarrow H^{-2}(\Omega)$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following stiffness and mass linear operators:
	\begin{align*}
	\left(\mathcal{K}u\right)(v) :=& \int_\Omega \nabla u \cdot \nabla v ~dx \\
	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
	\end{align*}
	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
	\begin{equation*}
	(\mathcal{P}u)_i = u(x_i)
	\end{equation*}
	for $i=1,\dots,r$.
\end{defn}

\begin{prop}
	\label{prop:equivalent_writing_of_weighting_optimization}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation*}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation*}
	may be rewritten as follows:
	\begin{equation}
		\label{eq:generic_optimization}
		\begin{aligned}
		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
		\text{such that} &\quad 
		\mathcal{P} u = f.
		\end{aligned}
	\end{equation}
\end{prop}

\begin{prop}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}. The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
	\begin{equation}
	\label{eq:weighting_kkt}
		\begin{bmatrix}
			\mathcal{K}\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
			\mathcal{P} & 0
		\end{bmatrix}
		\begin{bmatrix}
			u^* \\ \lambda^*
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ f
		\end{bmatrix},
	\end{equation}
	where we use block-matrix notation to denote block-operators.
\end{prop}

\begin{cor}
	\label{cor:linearity_of_solution_operator}
	Let $\mathcal{S}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
	\begin{equation*}
		\mathcal{S}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{S}$ is a linear operator.
\end{cor}

\begin{thm}
	\label{thm:smoothest_interpolant}
Here let $u$ be the function
\begin{equation*}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation*}
for an arbitrary but fixed value of $y$, where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{S}(f) &= \varphi_1(y) \mathcal{S}(\delta_{1j}) + \varphi_2(y) \mathcal{S}(\delta_{2j}) + \dots +\varphi_r(y) \mathcal{S}(\delta_{rj}) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u
	\end{align*}
 	because of linearity of the solution map, $\mathcal{S}$.
\end{proof}

\begin{thm}
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{thm}

\begin{proof}
asdf
\end{proof}

@@

@@

\section{Discretization}
\label{sec:discretization}

\subsection{Discretized operations}
\label{app:discretized_operations}

@@

\subsection{Discretized mean and covariance estimation} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively. The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
\end{align*}
respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.

\subsection{Discretized sampling of impulse responses}
In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$. The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$. The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
\begin{equation*}
\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
\end{equation*}
The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
\begin{equation*}
\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
\end{equation*}
